%% Bilingual Medical Interpreter CV – Sergio Muñoz de Alba Medrano
%% One-page optimized for Datamark, Inc. position
%% Compile with: pdflatex cv_Bilingual_Medical_Interpreter_ONE_PAGE.tex

\documentclass[8pt,letterpaper,sans]{moderncv}

% Style
\moderncvstyle{banking}
\moderncvcolor{blue}

% Encoding and fonts
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{lmodern}

% Ultra-compact one-page geometry
\usepackage[scale=0.62,footskip=14pt,top=0.25in,bottom=0.25in,left=0.35in,right=0.35in]{geometry}
\setlength{\hintscolumnwidth}{1.6cm}

% Smaller fonts
\renewcommand*{\namefont}{\fontsize{16}{18}\mdseries\upshape}
\renewcommand*{\titlefont}{\fontsize{8.5}{9}\mdseries\upshape}

% Ultra-compact spacing
\setlength{\parskip}{-3pt}
\setlength{\itemsep}{-5pt}
\setlength{\parsep}{-4pt}
\setlength{\topskip}{-2pt}
\setlength{\parindent}{0pt}
\setlength{\sectionsep}{-12pt}

% Personal data
\name{Sergio}{Muñoz de Alba Medrano}
\title{Doctor of Veterinary Medicine}
\address{Mérida, Yucatán}{Mexico}{}
\email{smunozam@gmail.com}
\phone[mobile]{+52~999~200~5550}
\extrainfo{LinkedIn: tinyurl.com/linkedinSMAM}

\begin{document}
\makecvtitle
\vspace{-10pt}
%----------------------------------------------------------------------------------------
% PROFESSIONAL SUMMARY
%----------------------------------------------------------------------------------------

\section{Professional Summary}

Bilingual Spanish–English professional with 20+ years of experience communicating technical information to diverse audiences. Extensive background in medical and veterinary terminology, regulatory systems, and training. Proven expertise in active listening, explaining complex concepts accessibly, and maintaining strict confidentiality. Seeking to provide professional interpretation services remotely.

%----------------------------------------------------------------------------------------
% CORE COMPETENCIES
%----------------------------------------------------------------------------------------

\section{Core Competencies}

\cvitem{Bilingual Communication}{Fluent spoken and written Spanish and English; professional communication, clear pronunciation, adaptability to different dialects and accents.}
\cvitem{Medical \& Veterinary Terminology}{Comprehensive knowledge of diagnosis, treatments, procedures, prescriptions, lab results, and animal health systems.}
\cvitem{Listening \& Comprehension}{Active listening, note-taking, memory retention, and ability to communicate with accuracy in fast-paced environments.}
\cvitem{Confidentiality \& Ethics}{Strict adherence to confidentiality protocols and professional ethical standards when handling sensitive health information.}

%----------------------------------------------------------------------------------------
% RELEVANT EXPERIENCE
%----------------------------------------------------------------------------------------

\section{Professional Experience}

\cventry{05/2025--12/2025}{Independent Consultant – Animal Health \& Governance}{SADER Yucatán}{Mérida}{}{
\begin{itemize}
\item Translated technical regulations between federal, state, and international stakeholders; managed bilingual documentation with strict accuracy and confidentiality.
\item Structured complex veterinary and sanitary information for diverse audiences; served as liaison between Spanish and English-speaking technical experts.
\end{itemize}}

\cventry{09/2017--12/2025}{Professor – Natural Resources \& Animal Production}{Universidad Marista de Mérida}{Mérida}{}{
\begin{itemize}
\item Taught animal production and sustainable agriculture using Spanish and English technical terminology.
\item Translated and explained complex concepts in simple language; adapted to different learning levels and developed mentoring skills.
\end{itemize}}

\cventry{10/2013--08/2017}{Agricultural Economist}{The Nature Conservancy}{Mérida}{}{
\begin{itemize}
\item Coordinated multidisciplinary teams, translating specialized concepts between technical experts and communities with varying educational backgrounds.
\item Maintained confidentiality of sensitive program information in all stakeholder interactions.
\end{itemize}}

%----------------------------------------------------------------------------------------
% EDUCATION, LANGUAGES & EQUIPMENT
%----------------------------------------------------------------------------------------

\section{Education}

\cvitem{1984--1988}{Bachelor's in Veterinary Medicine \& Animal Science, UNAM (Mexico City); Professional License}

\section{Languages}

\cvitem{Spanish}{Native proficiency}
\cvitem{English}{Bilingual proficiency (professional oral and written communication); TOEFL 670/677 (Paper-Based, Feb 1996)}

\section{Equipment \& Availability}

\cvitem{}{Computer and smartphone with stable internet; proficiency with videoconferencing platforms; available for remote work and flexible schedules.}

\end{document}
