%% Professional CV - Sergio Muñoz de Alba Medrano, DVM
%% English Version - January 2026 - 3 Page Optimized
%% Compile with: pdflatex peplexity_3_page_english.tex

\documentclass[10pt,letterpaper,sans]{moderncv}

% moderncv style
\moderncvstyle{banking}
\moderncvcolor{blue}

% Encoding and fonts
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{lmodern}

% Margin adjustment - tighter margins for 3 pages
\usepackage[scale=0.80,footskip=35pt]{geometry}
\setlength{\hintscolumnwidth}{2.2cm}

% Reduce name and title font size
\renewcommand*{\namefont}{\fontsize{18}{20}\mdseries\upshape}
\renewcommand*{\titlefont}{\fontsize{12}{14}\mdseries\upshape}

% Reduce spacing
\setlength{\parskip}{2pt}

% Personal data
\name{Sergio}{Muñoz de Alba Medrano}
\title{Doctor of Veterinary Medicine}
\address{Mérida, Yucatán}{Mexico}{}
\email{smunozam@gmail.com}
\phone[mobile]{+52~999~200~5550}
\homepage{tinyurl.com/linkedinSMAM}
\extrainfo{Operations \& Program Management | Chief of Staff}

%----------------------------------------------------------------------------------------
\begin{document}

\makecvtitle

%----------------------------------------------------------------------------------------
% PROFESSIONAL SUMMARY
%----------------------------------------------------------------------------------------

\section{Professional Summary}

\cvitem{}{Operations leader with \textbf{30+ years of experience} managing large-scale public programs, multidisciplinary teams (320+ staff), and budgets of \textbf{250+ million MXN}. Acts as \textbf{Chief of Staff partner} to senior leadership, aligning strategy with operations. Expertise in program execution, organizational development, process design, and stakeholder management across government, non-profit, and private sectors.}

%----------------------------------------------------------------------------------------
% CORE COMPETENCIES
%----------------------------------------------------------------------------------------

\section{Core Competencies}

\cvitem{Operations \& Leadership}{Multidisciplinary teams (320+ staff), 22 regional offices, clear objectives and performance monitoring.}
\cvitem{Program Management}{Multi-program portfolios: 250+ million MXN budgets, 45,000+ beneficiaries, strategic alignment.}
\cvitem{Financial Stewardship}{Budget oversight, resource allocation, audit support, transparency in public/non-profit settings.}
\cvitem{Org. Development}{Councils, communities of practice, cross-functional teams for collaboration and governance.}
\cvitem{Process \& Quality}{SOPs design, ISO 9000 participation, workflow optimization and continuous improvement.}
\cvitem{Stakeholder Relations}{Federal/state governments, international agencies, producers, NGOs, private sector partners.}
\cvitem{Digital Systems}{Web platforms, dashboards, data tracking systems (GOB.mx v3, bilateral agreement systems).}

%----------------------------------------------------------------------------------------
% PROFESSIONAL EXPERIENCE
%----------------------------------------------------------------------------------------

\section{Professional Experience}

\cventry{05/2025--Present}{Independent Consultant}{SADER Yucatán}{Mérida}{}{Strengthening operational systems, data management, and governance for animal health agreements. Led design and implementation of \textbf{Centro de Consulta de Acuerdos Zoosanitarios} (bilateral tracking system). Coordinated federal, state, and international stakeholders on procedures and compliance.}

\cventry{08/2020--04/2025}{Project Manager}{Youcatan Land}{Mérida}{}{Strategic projects in conservation and rural development. Secured state concession for \textbf{Tabi, Santuario de Selva Maya} (1,350 ha). Designed Integrated Strategies Program for agroecological transition with local stakeholders.}

\cventry{09/2017--12/2025}{Professor}{Universidad Marista de Mérida}{Mérida}{}{Faculty member in the Natural Resources Management program. Taught animal production systems and sustainable livestock design. Advised students on applied projects linking technical concepts with operational and organizational challenges.}

\cventry{10/2013--08/2017}{Agricultural Economist}{The Nature Conservancy}{Mérida}{M-REDD+}{Implemented silvopastoral systems in 8 pilot ranches. Served as Executive Secretary of the Sustainable Cattle Production Community of Practice. Coordinated technical teams, producers, and partner institutions, and contributed to reporting and knowledge management.}

\cventry{08/2012--02/2013}{Acting Head of Delegation}{SAGARPA}{Yucatán}{}{Highest federal agricultural authority in the state. Provided strategic and operational leadership to 320+ staff across 22 regional offices. Coordinated all federal agricultural program operations in Yucatán and led institutional relations with the State Government and sector stakeholders.}

\cventry{10/2005--09/2013}{Subdelegate for Planning}{SAGARPA}{Yucatán}{}{Operations and planning lead for federal agricultural programs in Yucatán. Managed the Procampo program (120M MXN annually to 45,000 producers). Oversaw execution of 250+ million MXN annually in federal–state concurrent programs. Implemented control mechanisms for legal compliance, fiscal discipline, and transparency, and led strategic planning processes for rural development.}

\cventry{08/2002--09/2005}{Head of Rural Development District}{SAGARPA}{Yucatán}{Ticul}{Regional operations lead for federal rural development programs. Coordinated subsidy programs for 18,000+ producers annually. Managed a team of 148 public servants and established 18 Municipal Councils for Sustainable Rural Development.}

\cventry{08/1998--07/2001}{Owner}{AGROPHARMA}{Yucatán}{}{Entrepreneur and general manager of a veterinary pharmaceutical microenterprise. Increased annual sales from 150,000 to 2 million MXN. Launched 15 new products and managed supplier relationships, inventory, quality controls, and administrative and commercial processes.}

\cventry{01/1994--12/1997}{New Product \& Export Manager}{LAPISA}{Michoacán}{}{Joint responsibilities in new product development and international export operations. Designed and established the export department from scratch. Initiated commercial operations in 13 countries in Latin America and Asia. Led preparation of regulatory dossiers and liaison with regulatory agencies, and participated in ISO 9000 certification processes.}

%----------------------------------------------------------------------------------------
% EDUCATION AND TECHNICAL SKILLS
%----------------------------------------------------------------------------------------

\section{Education}

\cventry{2004--2007}{Master's in Government and Public Policy (coursework completed)}{Universidad Autónoma de Yucatán}{Mérida}{}{}
\cventry{1994--1997}{Master's in Business Administration (coursework completed)}{Universidad del Valle de Atemajac}{La Piedad, Michoacán}{}{}
\cventry{1984--1988}{Bachelor's in Veterinary Medicine and Animal Science}{Universidad Nacional Autónoma de México}{Mexico City}{Professional License}{}

\section{Languages}

\cvitemwithcomment{Spanish}{Native}{}
\cvitemwithcomment{English}{Bilingual}{Professional oral and written communication}
\cvitemwithcomment{French}{Intermediate}{}

\section{Technical Skills}

\cvitem{Web \& App Development}{HTML, CSS, JavaScript, Vue.js, Firebase, GOB.mx v3 standards; development of small web applications for quality tools, financial calculations, and pension planning.}
\cvitem{Development Workflow}{Visual Studio Code, Git \& GitHub (version control, repositories for web apps and utilities), basic use of Node.js tooling.}
\cvitem{Data \& Analytics}{Data analysis, dashboards, basic use of statistical and geospatial tools (QGIS), integration of program data for reporting and decision-making.}
\cvitem{Productivity Tools}{Microsoft Office, Google Workspace; collaborative document management and presentations for senior stakeholders.}
\cvitem{AI \& Automation}{Use of Artificial Intelligence tools (including GitHub Copilot and other assistants) to draft documents, structure information, and accelerate analysis and reporting tasks.}

\end{document}
